---
title: "Instagram"
author: "Sahana Rayan"
date: "July 25, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
```

```{r echo = FALSE}
instagram_post_data <- read.csv("C:\\Users\\sahan\\Downloads\\dataset_my-task_2020-07-08_16-55-57-529.csv")
instagram_post_data_clean <- instagram_post_data %>% 
  separate(timestamp, c("Date", "Time"), "T")
```

#1) Are sidecars more interactive than just an image post.
- Side cars are the posts with a slide show of images (more than 1) and image posts are the posts with just one image. Since Purdue AKPsi page didn't post many videos, video posts were removed from the analysis
- __Hypothesis: Since people have short attention spans and are typically lazy, people would be more interested in image posts than sidecars. And therefore, they would be more likely to like an image post than a side car post__
- The interactiveness of a post will be measured by likes and comments received

```{r}
#table(instagram_post_data$type)
instagram_post_data %>% 
  filter(type != "Video") %>% 
  ggplot(aes(x = type, y = likesCount)) +
  geom_violin(size = 1.2, color = "#3A5382") +
  geom_boxplot(width = 0.2, fill = "#2D68C4", color = "#3A5382", size = 1.2) +
  theme_bw() +
  xlab("Types of posts") + ylab("Number of likes") +
  labs(
    title = "Distribution of likes for sidecars and images",
    caption = "Which type of post is more interactive"
  ) +
  scale_y_continuous(breaks = c(25, 50, 75, 100, 125)) +
  theme (
    plot.title = element_text(hjust = "0.4")
  ) +
  # theme(
  #   legend.position = "none",
  #   plot.background = element_rect(fill = "#3A5382",
  #                                  colour = "black",
  #                                  size = 1),
  #   axis.text = element_text(color = "white"),
  #   axis.title = element_text(color = "white"),
  #   plot.title = element_text(color = "white", hjust = "0.4"),
  #   plot.caption = element_text(color = "white"),
  #   panel.grid.minor.x = element_blank(),
  #   panel.grid.major.x = element_blank(),
  #   panel.grid.minor.y = element_blank(),
  #   panel.grid.major.y = element_blank(),
  #   panel.background = element_rect(fill = "#3A5382")
  # ) +
  coord_flip()
```


- This graph shows that not only is the median of likes for sidecar posts higher that that of image posts, the mode of likes is also greater for sidecar posts as displayed in the violin plot. 

```{r}
instagram_post_data %>% 
  filter(type != "Video") %>% 
  ggplot(aes(x = type, y = commentsCount)) +
  geom_boxplot(width = 0.3, fill = "#2D68C4", color = "#3A5382", size = 1.0) +
  theme_bw() +
  xlab("Types of posts") + ylab("Number of comments") +
  labs(
    title = "Distribution of comments for sidecars and images",
    caption = "Which type of post is more interactive"
  ) +
  theme (
    plot.title = element_text(hjust = "0.4")
  ) +
  coord_flip()

```


- Curious to see whether a similar trend would be observed with comments, a bar plot was plotted and yet again, side cars' median comments is greater than the image's posts

- From both these graphs, it can be concluded that sidecars seem to receive more likes and comments in general.

#2) Does the number of hashtags used for a post correlate to the number of likes
- I see a lot of public accounts use hashtags for the post to appear in non-follower's feeds. I am interested to see how effective this strategy is
- __Hypothesis : There will be a positive relationship between the number of hashtags and the number of likes till a certain point. After this saturation point, the number of likes will slowly decrease__


```{r} 
#This code helps count the hashtags
hashtagcols <- c("hashtags.0", "hashtags.1", "hashtags.2", "hashtags.3", "hashtags.4", "hashtags.5", "hashtags.6", 
                 "hashtags.7", "hashtags.8", "hashtags.9", "hashtags.10", "hashtags.11", "hashtags.12", "hashtags.13",
                "hashtags.14", "hashtags.15", "hashtags.16", "hashtags.17", "hashtags.18", "hashtags.19", "hashtags.20",
                 "hashtags.21", "hashtags.22", "hashtags.23", "hashtags.24", "hashtags.25", "hashtags.26","hashtags.27",
                 "hashtags.28", "hashtags.29") 

for (row in 1:nrow(instagram_post_data)) {
    count <- 0
    for (value in hashtagcols) {
      if (instagram_post_data[row, value] != "") {
        count <- count + 1
      }
    }
    instagram_post_data[row, "hashtagCount"] = count
  }
```

- A scatter plot can be useful in understanding the relationship between number of hashtags in a post and the number if likes for that post.
```{r warning = FALSE, message = FALSE}
instagram_post_data %>% 
  filter(hashtagCount > 0 ) %>% 
  ggplot(aes(x = hashtagCount, y = likesCount)) +
  geom_jitter(color = "#5C2F8C") +
  geom_smooth(width = 1.2, fill = "#EAAA00", color = "#ED8B00") +
  xlab("Number of hashtags") + ylab("Number of likes per post") +
  labs(title = "Scatter plot of number of hashtags and number of likes per post",
       caption = "What is the optimal number of hashtags")  +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = "0.5")
  )
  

```


- Although the line of best fit doesn't capture all the observations properly,  the number of likes appears to peak at a certain point ( 20 hashtags) after which there is a decline in the number of likes.
- This is what the distributions of hashtagcount variable is

```{r message = FALSE}
instagram_post_data %>% 
  ggplot(aes(x = hashtagCount)) +
  geom_histogram(fill = "#EAAA00", 
                 color = "#ED8B00",
                 size = 1.2) +
  geom_density() +
  xlab("Number of hashtags per post") + ylab("Number of posts") +
  labs(
    title = "Histogram showing the distribution of hashtags per post among all posts",
    caption = "What is the optimal number of hashtags?") +
  theme_minimal() +
  theme (
    plot.title = element_text(hjust = "0.5")
  )
```

- This shows that most of our posts have less than 10 hashtags.

- This distribution is used to keep only hashtagcounts with over 6 posts to remove outliers
```{r message = FALSE}
instagram_post_data %>%  
  group_by(hashtagCount) %>% 
  summarize(meanLikes = mean(likesCount), count = n()) %>% 
  filter (count >= 6) %>% 
  ggplot(aes(x = as.factor(hashtagCount))) +
  geom_bar(aes(y = meanLikes),
           stat = "identity", fill = "#EAAA00", color = "#ED8B00",
           size = 1.2) +
  xlab("Number of Hashtags") + ylab("Average likes") +
  labs(title = "Bar Chart of Average Likes for posts of different number of hashtags",
       caption = "What is the optimal number of hashtags") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = "0.5")
  )
```

- From this graph, it can be understood that there is a high correlation betweeen the number of hashtags and number of likes on a post, at least up until 18 hashtags.
- __NOTE: Correlation does not mean causation. I am, in no way, a causal inference expert deduce anything like that. But, I can analyze patterns.__
- 18 hashatgs also appears to be the optimal number of hashtags




#3) What are the most effective hashtags then?
- The last research question led me to wonder what hashtags are effective in bringing likes and comments. What hashtags bring the most public attention. 
- __Hypothesis : The less specific hashtags like purdue and university are the most effective__

```{r}
hashtag_data <- instagram_post_data %>% 
  pivot_longer(hashtagcols,
               names_to = "HashtagNo",
               values_to = "HashtagDesc") %>% 
  filter(HashtagDesc != "") %>% 
  mutate(likeContrib = likesCount/ hashtagCount)

hashtag_data %>% 
  group_by(HashtagDesc) %>% 
  summarize(avg_likes = mean(likesCount), count = n()) %>% 
  filter(count >= 16) %>% 
  ggplot(aes(x = HashtagDesc)) +
  geom_bar(aes(y = avg_likes), 
           stat = "identity", fill = "#2D68C4", color = "#3A5382", size = 1.2) +
  labs(
    title = "Bar Chart of Avg likes for posts with this selection of hashtags",
    caption = "What are the most effective hashtags then?",
    x = "Hashtag Name",
    y = "Average likes per post"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = "0.4"),
    axis.text.x = element_text(angle = 45, vjust = "0.5")
  )

```

- From the looks of this graph, hashtags with appropriate capitalized letters have higher average likes
- However, these results might be biased by posts with many hashtags. To remove this potential factor (number of hashtags) from the analysis,  I looked at the avg likes per hashtag or like contribution that each hashtag brings instead of looking at the average likes per post.
-So if a post received 50 likes and used 10 hashtags, the like contribution for each hashtag in the post would be 5 likes
- __NOTE: I am assuming in this analysis that each hashtag of a post has an equal contribution to the likes of the post__
- A more appropiate analysis of like contribution can be done from ridge plots that will show the distribution of like contribution for each type of hashtage. This plot only displays hashtags that appear in more than 15 posts. This subsetting ensures a lack of bias 
```{r echo = FALSE, fig.show='hide'}
hashtag_data %>% 
  group_by(HashtagDesc) %>% 
  summarize(avg_likesContrib = mean(likeContrib), count = n()) %>% 
  filter(count > 15) %>% 
  ggplot(aes(x = HashtagDesc)) +
  geom_bar(aes(y = avg_likesContrib), 
           stat = "identity",
           fill = "#2D68C4", color = "#3A5382", size = 1.2) +
  labs(
    title = "Bar Chart of Avg like contribution for this selection of hashtags",
    caption = "What are the most effective hashtags then?",
    x = "Hashtag Name",
    y = "Average like contribution"
  ) +
  theme_classic()  +
  theme(
    plot.title = element_text(hjust = "0.4"),
    axis.text.x = element_text(angle = 45, vjust = "0.5")
  )
  

```


```{r echo= FALSE}
tt <- table(hashtag_data$HashtagDesc)

```

```{r}
library(ggridges)
hashtag_data[hashtag_data$HashtagDesc %in% names(tt[tt >15]),] %>% 
  ggplot(aes(x = likeContrib, y = HashtagDesc)) +
  geom_density_ridges(rel_min_height = 0.01, fill = "#2D68C4", color = "#3A5382", size = 1.0) +
  labs(
    title = "Ridge Plots of like contribution for this selection of hashtags",
    caption = "What are the most effective hashtags then?",
    x = "Like contribution",
    y = "Hashtag Name"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = "0.4")
  )
```


- It can be derived from the plot that Purdue-related hashtags have very similar distributions with mode at around 5 likes (contribution)
- __akpsi__, __university__, and __piomega__ have different distributions than that of the the purdue-related ones.
- __university__'s distribution is bimodal but it's actual mode seems to be at around 9 likes (contribution)
- Both __akpsi__ and __piomega__ have modes at around 7.



#4) An optimal day of the week?


```{r}
library(stringr)
library(lubridate)
library(ggbeeswarm)
instagram_post_data_clean$Date <- as.Date(instagram_post_data_clean$Date)
instagram_post_data_clean %>% 
  mutate(wday = weekdays(Date)) %>% 
  ggplot(aes(x = wday, y = likesCount)) +
  geom_boxplot(fill = "#EAAA00", color = "#ED8B00",
           size = 1.0, alpha = 0.8) +
  labs(
    x = "Number of Likes",
    y = "Day of the week",
    title = "Distribution of likes for different days of the week",
    caption = "What is the best day of the week to post on?"
  ) +
  scale_x_discrete(limits = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) +
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125)) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = "0.4")
  )


```

- From this graph, we can tell that posts published on Sundays and Saturdays to bring the least likes
- But, posts from Fridays, Thursday or in the middle of the week produced more likes in general

#5) Time Series Analysis
- This time series plot shows some of the trends that we have seen over time
- The trend line uses a generalized additive model to approximate the general trend
```{r}
instagram_post_data_clean %>% 
  ggplot(aes(x = Date, y = likesCount)) +
  geom_point(color = "#5C2F8C") +
  geom_smooth(fill = "#2D68C4", color = "#3A5382", size = 1.0, method = "gam") +
  #geom_line(color = "#3A5382", size = 1.0) +
  labs(
    x = "Number of likes",
    y = "Timestamp",
    title = "Likes-Time Plot",
    caption = "Time Series Analysis"
  ) +
  theme_minimal() +
  scale_x_date(date_breaks = "6 month", date_labels = "%b %Y") +
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125)) +
  theme(
    plot.title = element_text(hjust = "0.5"),
    axis.text.x = element_text(angle = 45),
    axis.title.x = element_text(vjust = "0.5")
  )

```

- There is an overall positive trend in the number of likes received on each post
- There seems to be a dip in likes between July 2018 and June 2019

